\Large\textbf{Advection versus diffusion in brain ventricular transport}

### 2025-04-07 paper/main.tex
Updates in Discussion

\subsection{Effects of cilia motility on transport}
Removing motile cilia slowed down transport within
the middle ventricle and towards the anterior ventricle,
but the transport to the posterior ventricle was unaffected.
Photoconversion experiments with cilia mutant
zebrafish larva (\emph{smh}) aligned with these observations.
This further indicates that
transport from the middle to the posterior ventricle is
purely diffusive at this developmental stage,
while transport within the middle ventricle
depends on cilia motility.
Interestingly, Yoshida~\emph{et al.} observed a drastic reduction
in CSF mixing in specific regions of the brain
ventricles when simulating reduced cilia motility~\cite{Yoshida2022EffectVentricles}.
Our experimental results provide not only confidence
in the predictive ability of our computational model,
but also quantify the impact of cilia motility
on brain ventricular transport.

\subsection{Model parameters and sources of discrepancies}
Even though we identified similar dynamics in simulations and
in vivo experiments, the transport dynamics of the simulations were generally
more rapid and resulted in higher total amounts of solute.
Differences between experiments and simulations may relate to
ventricular size, as we showed that modifying the ventricular geometry impacts solute distribution,
and the ventricular size varies substantially 
in the control population at 2 dpf.
Additionally, differences may result
from the definition of the photoconversion site,
since the volume of this region determines the total amount of solute
and thereby the rate of transport. Another source
of discrepancy may relate to the permeability of the ventricular wall,
which we considered as impermeable in our simulations.
The neuroepithelium lining the walls of brain ventricles in 24 hour old zebrafish
is permeable to dyes smaller than 70 kDa~\cite{Chang2012AnNeuroepithelium}.
If this is also the case at 2 dpf, Dendra2 
(which are 26 kDa in size~\cite{Gurskaya2006EngineeringLight})
would leak out of the brain ventricles over the relevant time spans
and render the dynamics slower.
It may also be the case that our model predicts faster
transport because the Dendra2 diffusion
coefficient we used (based on an experimental value
\cite{GuraSadovsky2017MeasurementExpansion}) is too high.
This could be because embryonic CSF has higher viscosity
than the solution considered in the measurements, 
or because the presence of proteins, nutrients and other solutes
obstructs diffusive transport
\cite{Bloomfield1998EffectsFluid, Einstein1905UberTeilchen, Cussler2009Diffusion:Systems}.
Finally, an important difference
is the three-dimensional nature of the computational model,
while the experimental images only show 2D slices.

@@

\subsection{Numerical approximation}
When solving an advection-diffusion equation to
simulate protein transport, we handled advective transport
through the anterior/posterior boundaries
(where cardiac pressure pulsations are imposed)
with an approximate periodic boundary condition
that conserves the total solute concentration.
A possible improvement to this would be
to consider a 3D-0D coupling of in- and outflow using
ordinary differential equations,
similar to applications in cardiac
modeling~\cite{Brown2024AMechanics, Augustin2021ACirculation}.
Alternatively, the beating cardiac motion could be modeled by ventricular wall deformation
to avoid advective boundary fluxes\cite{Causemann2022HumanFramework,
Kurtcuoglu2005ComputationalSystem, Kurtcuoglu2007ComputationalSylvius, Linninger2005PulsatileBrain}.

The Brezzi-Douglas-Marini and discontinuous Galerkin scheme
$\mathrm{BDM}_1$-$\mathrm{DG}_0$
we used to discretize the Stokes equations
is computationally more expensive
than \fixme{conventional methods} \miro{ the more commonly used discretizations} such as a
Taylor-Hood ($P_2-P_{1}$ elements)~\cite{Stenberg1990ErrorProblem} or lower-order enriched
elements (e.g. MINI)~\cite{Brezzi2011MixedMethods}.
Here, $P_k$ is the space of continuous Lagrange polynomials of degree less
than or equal to the integer $k$.
We opted for the BDM-DG scheme because of its exactly-mass-conserving property;
by the definition of the elements, the numerical approximation of
the CSF velocity $\uu_h$ satisfies the continuity equation (\Cref{eq:stokes_eqs_cont}).
As a result $\normltwo{\nabla\cdot\uu_h}=0$ to within solver tolerances,
an important property for the stability of the advection-diffusion
equation, especially in advection-dominated cases.
\cite{cesmelioglu2022compatible, johnson2009numerical, gresho1998incompressible}
Additionally, the BDM-DG scheme 
lets us impose the impermeability condition $\uu\cdot\nn=0$
strongly, since the BDM elements control the normal component of the 
velocity~\cite{Brezzi1985TwoProblems}.
For elements where this is not the case (such as Taylor-Hood elements),
the impermeability boundary condition must
be handled with for example a Nitsche method~\cite{Nitsche1971UberSind}
or with a Lagrange multiplier~\cite{Babuska1973TheMultipliers,
Bertoluzza2017BoundaryHemodynamics}, which affects
convergence properties \miro{$<<$what do you mean here?} and increases problem complexity
(see \Cref{subsec:appendixA1} for more details).

\subsection{Outlook}
The flow generated by ciliary motion
in zebrafish larval brain ventricles
is of the scale of the geometry~\cite{Olstad2019CiliaryDevelopment},
while in large brains such as the human brain, ciliary beating 
contributes to flow proximal to the ventricular walls
\cite{Siyahhan2014FlowVentricles}.
The cilia may play an important role in regulating the
amount of substances like extracellular vesicles close to the ventricular walls,
similar to coral reefs where cilia are primary regulators of
oxygen amounts at the coral surface~\cite{Pacherres2022CiliaryProduction}.
Cilia indeed regulate nodal flow at early
developmental stages of the brain~\cite{Hirokawa2006NodalAsymmetry},
and the discovery of a cilia-driven flow network
in rodent brains suggests such selective
brain ventricular transport~\cite{Faubel2016Cilia-basedVentricles}.
As the CSF flows through the central nervous system,
the system may be affected by the changes in CSF composition
due to reduced mixing following reduced
ciliary motion. We demonstrated with
simulations that solute concentrations,
and in turn CSF composition, 
are affected by ventricular
geometry changes. A 10\% difference in volume
had a noticeable impact on solute transport,
showing that pathological ventricular volume changes,
such as the enlargement observed in hydrocephalus,
would affect CSF composition.
Because CSF plays an important role in the regulation of solutes
within the ventricular system~\cite{yamada2021cerebrospinal},
understanding more of the interplay between
transport dynamics and ventricular morphology,
as well as the impact of
reduced mixing resulting from cilia-deficiency,
are important considerations for future work.
We hope our findings and the computational
framework we have presented can help address
such open questions in the future.

>>>>

\subsection{Effects of cilia motility on transport}

Removing motile cilia delayed transport within
the middle ventricle and towards the anterior ventricle,
while the transport to the posterior ventricle was unaffected.
Photoconversion experiments with cilia mutant
zebrafish larva (\emph{smh}) aligned with these observations.
This further indicates that
transport from the middle to the posterior ventricle is
purely diffusive at this developmental stage,
while transport within the middle ventricle
depends on cilia motility.
Interestingly, Yoshida~\emph{et al.} observed a drastic reduction
in CSF mixing in specific regions of the brain
ventricles when simulating reduced cilia motility~\cite{Yoshida2022EffectVentricles}.
\fixme{Our experimental results provide not only confidence
in the predictive ability of our computational model,
but also quantify the impact of cilia motility
on brain ventricular transport.}

\subsection{Model validation}

Our computational predictions and in vivo experimental measurements
agree to a large extent. Examining any discrepancies in more detail
indicates that the simulated transport dynamics are more rapid and
results in higher total amounts of solute. Differences between
experiments and simulations may relate to ventricular size, as we
showed that modifying the ventricular geometry impacts solute
distribution, and the ventricular size varies substantially in the
control population at 2 dpf. In addition, differences may result from
the definition of the photoconversion site, since the volume of this
region determines the total amount of solute and thereby the rate of
transport. Another source of discrepancy may relate to the
permeability of the ventricular wall, which we considered as
impermeable in our simulations. The neuroepithelium lining the walls
of brain ventricles in 24 hour old zebrafish is permeable to dyes
smaller than 70 kDa~\cite{Chang2012AnNeuroepithelium}.  If this is
also the case at 2 dpf, Dendra2 (which are 26 kDa in
size~\cite{Gurskaya2006EngineeringLight}) would leak out of the brain
ventricles over the relevant time spans and render the dynamics
slower.  It may also be the case that our model predicts faster
transport because the Dendra2 diffusion coefficient we used (based on
an experimental value \cite{GuraSadovsky2017MeasurementExpansion}) is
too high. This could be because embryonic CSF has higher viscosity
than the solution considered in the measurements, or because the
presence of proteins, nutrients and other solutes obstructs diffusive
transport \cite{Bloomfield1998EffectsFluid, Einstein1905UberTeilchen,
  Cussler2009Diffusion:Systems}. Finally, an important difference is
the three-dimensional nature of the computational model, while the
experimental images only show 2D slices.

@@

\subsection{Numerical considerations}
When simulating solute transport, we handled the advective transport
through the anterior/posterior boundaries
(where cardiac pressure pulsations were imposed)
via an approximate periodic boundary condition
that conserved the total solute concentration.
A possible improvement would be
to consider a 3D-0D coupling of the in- and outflow using
ordinary differential equations,
similar to applications in cardiac
modeling~\cite{Brown2024AMechanics, Augustin2021ACirculation}.
Alternatively, the beating cardiac motion could be modeled by ventricular wall deformation to avoid advective boundary fluxes\cite{Causemann2022HumanFramework,
Kurtcuoglu2005ComputationalSystem, Kurtcuoglu2007ComputationalSylvius, Linninger2005PulsatileBrain}.
%
We also remark that the Brezzi-Douglas-Marini and discontinuous Galerkin scheme
$\mathrm{BDM}_1$-$\mathrm{DG}_0$
used to discretize the Stokes equations
is computationally more expensive
than more commons discretizations such as 
Taylor-Hood ($P_2-P_{1}$ elements)~\cite{Stenberg1990ErrorProblem} or lower-order enriched
elements (e.g. MINI)~\cite{Brezzi2011MixedMethods}.
%Here, $P_k$ is the space of continuous Lagrange polynomials of degree less
%than or equal to the integer $k$.
However, we prefer the BDM-DG scheme because of its mass conservation properties: the numerical approximation of
the CSF velocity $\uu_h$ satisfies the continuity equation (\Cref{eq:stokes_eqs_cont}) point-wise, an important property for the stability of the advection-diffusion
equation, especially in advection-dominated cases.
\cite{cesmelioglu2022compatible, johnson2009numerical, gresho1998incompressible}
Additionally, the BDM-DG scheme 
lets us impose the impermeability condition $\uu\cdot\nn=0$
directly~\cite{Brezzi1985TwoProblems} (see \Cref{subsec:appendixA1} for more details).
%% For elements where this is not the case (such as Taylor-Hood elements),
%% the impermeability boundary condition must
%% be handled with for example a Nitsche method~\cite{Nitsche1971UberSind}
%% or with a Lagrange multiplier~\cite{Babuska1973TheMultipliers,
%% Bertoluzza2017BoundaryHemodynamics}, which affects
%% convergence properties \miro{$<<$what do you mean here?} and increases problem complexity
%% (see \Cref{subsec:appendixA1} for more details).

\subsection{Outlook}
The flow generated by ciliary motion
in zebrafish larval brain ventricles
is of the scale of the geometry~\cite{Olstad2019CiliaryDevelopment},
while in large brains such as the human brain, ciliary beating 
contributes to flow proximal to the ventricular walls
\cite{Siyahhan2014FlowVentricles}.
The cilia may play an important role in regulating the
amount of substances like extracellular vesicles close to the ventricular walls,
similar to coral reefs where cilia are primary regulators of
oxygen amounts at the coral surface~\cite{Pacherres2022CiliaryProduction}.
Cilia indeed regulate nodal flow at early
developmental stages of the brain~\cite{Hirokawa2006NodalAsymmetry},
and the discovery of a cilia-driven flow network
in rodent brains suggests such selective
brain ventricular transport~\cite{Faubel2016Cilia-basedVentricles}.
As the CSF flows through the central nervous system,
the system may be affected by the changes in CSF composition
due to reduced mixing following reduced
ciliary motion. We demonstrate that solute concentrations,
and in turn CSF composition, 
are affected by ventricular
geometry changes. A 10\% difference in volume
had a noticeable impact on solute transport,
showing that pathological ventricular volume changes,
such as the enlargement observed in hydrocephalus,
could affect CSF composition.
Because CSF plays an important role in the regulation of solutes
within the ventricular system~\cite{yamada2021cerebrospinal},
understanding more of the interplay between
transport dynamics and ventricular morphology,
as well as the impact of
reduced mixing resulting from cilia-deficiency,
are important considerations for future work.
%We hope our findings and the computational
%framework we have presented can help address
%such open questions in the future.
